---
title: "Zhanran L01"
subtitle: "Data Visualization (STAT 302)" 
author: "Zhanran Shi"

format:
  html:
    toc: true
    embed-resources: true
    link-external-newwindow: true

execute:
  warning: false
  echo: false

from: markdown+emoji 
---


## Goals

1.  Ensure software for this course is properly installed and functional
2.  Understand the importance of visualization
3.  Develop and follow a proper workflow
4.  Construct and explore a few plots using `ggplot2`
5.  Add layers to a plot that include title, axis, and pre-set themes

**Don't worry if you cannot do everything here by yourself.** You are just getting started and the learning curve is steep, but remember that the instructional team and your classmates will be there to provide support. Persevere and put forth an honest effort and this course will payoff.


```{r}
#| label: load-packages

library(ggplot2)
library(tidyverse)
library(dplyr)
library(forcats)
load('/Users/zhanranshi/Desktop/2112/Junior/2024 Winter/STAT 302/Lec Submissions/Data/legosets.rda')
```


## Tasks

Complete the following tasks. For many of these you'll need to simply **indicate that you have completed the task**.

### Task 1

Download and install [R Software](https://cran.rstudio.com/).

### Task 2

1.  Download and install [RStudio](https://www.rstudio.com/products/rstudio/download4/).
2.  Open RStudio and ensure it and R have been properly installed.
3.  Go to **Tools** \> **Global Options** and change the two indicated settings below and click **Apply**. Your other options are likely different than what is pictured. This is okay and we suggest you leave them as is until you become more familiar with RStudio.

**See html file for missing figure**

### Task 3

Take some time to explore the [Quarto](https://quarto.org/) and [ggplot2](https://ggplot2.tidyverse.org/index.html) websites. We suggest bookmarking them because they will be extremely useful references.

We highly recommend reading through the [Tutorial: Hello, Quarto](https://quarto.org/docs/get-started/hello/rstudio.html).

### Task 4

Install the following packages:

-   `tidyverse`, actually it is a collection/universe of packages being installed
-   `skimr`
-   `devtools`

### Task 5

1.  Download `L01_introduction.zip` from Canvas.
2.  Unzip the file and place the `data_vis_labs` directory where you would like to keep all of your lab work for this course.
3.  Open RStudio and create a project folder for this *existing directory*.
4.  Rename `template_L01.qmd` to your `Lastname_Firstname_L01.qmd`
5.  Render the file with `Cmd/Ctrl + Shift + K` or click the Render button.

### Task 6 (optional)

If you are new to RStudio we strongly recommend reading through Chapters 1-4 of the [Stat 202 text book](https://nustat.github.io/intro-stat-data-sci/)

## Exercise 1

Let's delve into why data visualizations are important.

### Visualization 1

Read the following article and answer the following questions.

[bars_vs_grocery_stores](https://flowingdata.com/2014/05/29/bars-versus-grocery-stores-around-the-world/)

1)  List one thing you learned from the article?

That surprisingly grocery stores per capita is higher than bars per capita in most of Germany and Poland, whereas in Wisconsin, where residents from both countries make up large parts of the population, it's the opposite.

2)  How did the main visualization (United States) enhance the article?

Firstly it helped validate Floatingsheep's conclusion by showing the Wisconsin area in a much darker color than the rest of the country. Secondly it also helped reveal nuances in this census as in most bar areas the color was very light, showing how the bars did not outnumber grocery stores as strongly as Floatingsheep's more binary visualization would suggest.

3)  List at least one thing you liked about the \[United States\] graphic and one thing you think could be improved.

I liked how the visualization gave an extra dimension to the bar vs grocery ratio using color-coding. I wish the state borders could be more well-defined although I do understand the practical difficulties adding more elements might introduce to the visualization effectiveness.

<br>

### Visualization 2

Open this [dashboard](https://edgarruiz.shinyapps.io/db-dashboard/) and answer the following questions.

1)  What story is this dashboard telling and how do the visualizations enhance that?

The dashboard tells me the total number of flights, daily averages, delay statistics, monthly flight trends, and popular destinations. Monthly flight distribution and destination airport flights are visualized as bar charts, making it easy to compare values and spot patterns. The color difference also makes the most popular destination figure more impressionistic to a quick reader.

2.  What might a dashboard provide that a static graphic cannot.

Dashboards combine various visualizations and metrics, offering a holistic view of the data without needing multiple static charts. Dashboards also allow users to click on elements for more detailed data, filter views, or customize the scope of analysis, unlike static graphics

## Exercise 2

Let's look at some interesting patterns in the history of LEGO! We'll be using data from the [`lego` package](https://github.com/seankross/lego) located `data/legosets.rda`.

First, let's look at the average cost of LEGO sets over time. The main variable of interest here is `USD_MSRP`, or the manufacturer's suggested retail price in constant dollars.


```{r}

lego_avg <- legosets %>%
              group_by(Year) %>%
              summarize(avg_msrp = mean(USD_MSRP, na.rm = TRUE))
ggplot(lego_avg, aes(Year, avg_msrp)) + 
  geom_line() + 
  geom_point() +
  labs(
    title = "Average price of LEGO sets",
    subtitle = "Amounts are reported in current USD ",
    x = "Year",
    y = "Price (USD)",
    caption = "Source: LEGO"
    ) +
  theme_minimal()
```

## Exercise 3

Let's explore how many unique themes LEGO now offers.

```{r}
#| label: themes
 
# remove #| eval = FALSE so the plot appears 

theme_counts <- legosets %>%
  count(Theme, sort = TRUE) %>%
  mutate(Theme = fct_inorder(Theme, ordered = TRUE))

theme_counts %>%
  ggplot(aes(y = Theme, x = n)) +
  geom_col() +    # or geom_bar(stat = "identity") +
  labs(
    x = "Number of Sets",
    y = NULL
  ) +
  theme_minimal()
```

What is bad/ugly about the plot above?

Range is all over the place and the y labels cover each other.

Now remake the plot BUT with the improvements addressing the issues you noted above.

```{r}
theme_counts %>% 
  filter(n > 150) %>%
  ggplot(aes(x = n, y = fct_rev(Theme))) +
  geom_col() +
  labs(
    x = "Number of Sets",
    y = NULL,
    title = "Number of Sets per Theme"
  ) +
  theme_minimal()
```


## Exercise 4

Now you try, let's look at how the number of pieces per set has changed over time. Because Duplo sets are much smaller (since they're designed for toddlers), we made a special indicator variable for them shown in the code below.

Use an appropriate graph to plot the average number of pieces over time. Have either separate lines or separate plots for the Duplo and Lego categories. Be sure the title, x-axis, and y-axis are properly labeled.

```{r}
#| label: lego-pieces

# remove #| eval = FALSE or this chunk won't evaluate

pieces_per_year <- legosets %>%
  # new indicator variable
  mutate(Duplo = ifelse(Theme == "Duplo", "Duplo", "Lego")) %>%
  group_by(Year, Duplo) %>%
  summarize(avg_piece = mean(Pieces, na.rm=TRUE))
  # summarize() mean pieces

ggplot(pieces_per_year, aes(y = avg_piece, x = Year)) + 
  geom_line() +
  facet_wrap(~Duplo) + 
  geom_smooth() +
  labs(
    x = "Year",
    y = "Average Piece Count",
    title = "Average piece per year for Lego and Duplo"
  ) +
  theme_minimal()

```

What trends do you notice in the plots? Are you surprised by the plots?

I noticed that the average pieces of Duplo overall decreased and the average pieces of Lego overall increased. I'm surprised that there's a small rise in average pieces from mid-2000 to early 2010s but fell again, so I'd assume the more complex new sets weren't gaining popularity and so they reverted to simpler designs.
